---
title: Final Project
output: html_document
---


```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(readr) 
library("ggplot2") 
library("dplyr")
library(xts)
library("lubridate")
library("RColorBrewer")
library("ggthemes")
library(gridExtra)
library(corrplot)
library(fmsb)

data <- read_csv("./Project_1_data.csv")

for (column in colnames(data)) {
  num_distinct_values <- length(unique(data[[column]]))
  print(paste(column, ":", num_distinct_values, "distinct values"))
}

library(ggplot2)
```

```{r}
# Create a ggplot for the distribution of MathScore
plot_math <- ggplot(data, aes(x = MathScore)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  labs(title = "Distribution of MathScore", x = "Score")

# Create a ggplot for the distribution of ReadingScore
plot_reading <- ggplot(data, aes(x = ReadingScore)) +
  geom_histogram(fill = "lightcoral", color = "black", bins = 30) +
  labs(title = "Distribution of ReadingScore", x = "Score")

# Create a ggplot for the distribution of WritingScore
plot_writing <- ggplot(data, aes(x = WritingScore)) +
  geom_histogram(fill = "lightgreen", color = "black", bins = 30) +
  labs(title = "Distribution of WritingScore", x = "Score")

grid.arrange(plot_math, plot_reading, plot_writing, ncol = 3)

ggtitle("Distribution of the test scores")
```

To deal with NA values, we replaced every NA value with the most frequent value in each column. 
```{r}
data_cleaned <- apply(data, 2, function(x) ifelse(is.na(x), names(sort(table(x), decreasing = TRUE)[1]), x))

null_val <- sum(is.na(data_cleaned))

if (null_val > 0) {
  print(paste('Try imputing again, there are', null_val, 'null values remaining in the dataset'))
} else {
  print('Congratulations! You have successfully imputed all the null values in the dataset.')
}
```

```{r}
data_cleaned <- data.frame(data_cleaned)

order_parent_educ_categories <- c('some high school', 'high school', "associate's degree", 'some college', "bachelor's degree", "master's degree")

order_ethnic_categories <- c('group A', 'group B', 'group C', 'group D', 'group E')

study_categories <- c('< 5', '5-10', '> 10')

data_cleaned <- data_cleaned %>% 
  mutate(Gender = as.factor(Gender), 
         MathScore = as.integer(MathScore), 
         ReadingScore = as.integer(ReadingScore), 
         WritingScore = as.integer(WritingScore),
         ParentEduc = factor(ParentEduc, levels = order_parent_educ_categories, ordered = TRUE), 
         EthnicGroup = factor(EthnicGroup, levels = order_ethnic_categories, ordered = TRUE)) %>% 
  mutate(ln_math = log(MathScore), 
         ln_reading = log(ReadingScore), 
         ln_writing = log(WritingScore))

model_1 <- lm(ReadingScore ~ LunchType, data = data_cleaned)
summary(model_1)

model_2 <- lm(ReadingScore ~ EthnicGroup, data = data_cleaned)
summary(model_2)

model_3 <- lm(ReadingScore ~ ParentEduc, data = data_cleaned)
summary(model_3)

model_4 <- lm(ReadingScore ~ Gender, data = data_cleaned)
summary(model_4)

model_5 <- lm(ReadingScore ~ TestPrep, data = data_cleaned)
summary(model_5)

model_6 <- lm(ReadingScore ~ ParentMaritalStatus, data = data_cleaned)
summary(model_6)

model_7 <- lm(ReadingScore ~ PracticeSport, data = data_cleaned)
summary(model_7)

model_8 <- lm(ReadingScore ~ IsFirstChild, data = data_cleaned)
summary(model_8)

model_9 <- lm(ReadingScore ~ NrSiblings, data = data_cleaned)
summary(model_9)

model_10 <- lm(ReadingScore ~ WklyStudyHours, data = data_cleaned)
summary(model_10)

model_11 <- lm(ReadingScore ~ TransportMeans, data = data_cleaned)
summary(model_11)

model_1 <- lm(MathScore ~ LunchType, data = data_cleaned)
summary(model_1)

model_2 <- lm(MathScore ~ EthnicGroup, data = data_cleaned)
summary(model_2)

model_3 <- lm(MathScore ~ ParentEduc, data = data_cleaned)
summary(model_3)

model_4 <- lm(MathScore ~ Gender, data = data_cleaned)
summary(model_4)

model_5 <- lm(MathScore ~ TestPrep, data = data_cleaned)
summary(model_5)

model_6 <- lm(MathScore ~ ParentMaritalStatus, data = data_cleaned)
summary(model_6)

model_7 <- lm(MathScore ~ PracticeSport, data = data_cleaned)
summary(model_7)

model_8 <- lm(MathScore ~ IsFirstChild, data = data_cleaned)
summary(model_8)

model_9 <- lm(MathScore ~ NrSiblings, data = data_cleaned)
summary(model_9)

model_10 <- lm(MathScore ~ WklyStudyHours, data = data_cleaned)
summary(model_10)

model_11 <- lm(MathScore ~ TransportMeans, data = data_cleaned)
summary(model_11)
```

```{r}
# Create a ggplot for the distribution of MathScore
plot_math <- ggplot(data_cleaned, aes(x = ln_math)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 50) +
  labs(title = "Distribution of MathScore", x = "Score")

# Create a ggplot for the distribution of ReadingScore
plot_reading <- ggplot(data_cleaned, aes(x = ln_reading)) +
  geom_histogram(fill = "lightcoral", color = "black", bins = 50) +
  labs(title = "Distribution of ReadingScore", x = "Score")

# Create a ggplot for the distribution of WritingScore
plot_writing <- ggplot(data_cleaned, aes(x = ln_writing)) +
  geom_histogram(fill = "lightgreen", color = "black", bins = 50) +
  labs(title = "Distribution of WritingScore", x = "Score")

grid.arrange(plot_math, plot_reading, plot_writing, ncol = 3)

ggtitle("Distribution of the ln test scores")
```

Gender

- The distribution of male and female students in the dataset is fairly balanced.
- Score wise girls performed better that boys in ReadingScore andwritingScore, whereas boys scored better in MathScore.

```{r}
ggplot(data_cleaned, aes(x = Gender)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Gender") +
  theme_minimal()
```

```{r}
mean_scores <- data_cleaned %>%
  group_by(Gender) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )

print(mean_scores)
```

Parent(s) educational background

- There are six categories of educational background of parent(s) of the students, ranging from master's education (the highest level) to highschool dropouts (some high school)
- In terms of distribution, master's degree holder parent(s) are the least at 5.8% and parent(s) with some college are the most at 26.6%.
- When it comes to the students' test scores, there seems to be a clear correlation with their parents(s) educational background. Children of highly educated parents score better than the children of less educated parents. This pattern is seen consistently with all three test subjects.

```{r}
order_parent_educ_categories <- c('some high school', 'high school', "associate's degree", 'some college', "bachelor's degree", "master's degree")

ggplot(data_cleaned, aes(x = factor(ParentEduc, levels = order_parent_educ_categories))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Parents educational background") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
```
```{r}
bar_parent <- ggplot(data = data_cleaned) + 
  geom_bar(
    mapping = aes(x = ParentEduc, fill = ParentEduc), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar_parent + coord_polar()
```

```{r}
val_cnt <- as.data.frame(table(data_cleaned$ParentEduc))
val_cnt_norm <- prop.table(table(data_cleaned$ParentEduc)) * 100

val_cnt
val_cnt_norm <- as.data.frame(val_cnt_norm)
val_cnt_norm

mean_score_summary <- data_cleaned%>%
  group_by(ParentEduc) %>%
  summarise_all(mean) %>% 
  select(ParentEduc,MathScore, ReadingScore, WritingScore)

mean_score_summary
```

Ethnic group

- There are five races/ethnic groups represented in the dataset; (anonymized) from group A to group E
- With 35.4% of the student population, race group C are the most represented whereas with only 8.4% race group A are the least.
- The effect of race on students' performance is clear to see. Race group E students are by far the better performers having scored 6 points more than the second group (D) in math and 3 more points in reading score and close to 1 point in writing test. At the other end of the scoreboard we find race group A; this group scored almost 12 points on math lower than that of group E. 

```{r}
order_ethnic_categories <- c('group A', 'group B', 'group C', 'group D', 'group E')

ggplot(data_cleaned, aes(x = factor(EthnicGroup, levels = order_ethnic_categories))) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Which ethnic group do the students belong to?") +
  theme_minimal()
```
```{r}
bar_parent <- ggplot(data = data_cleaned) + 
  geom_bar(
    mapping = aes(x = EthnicGroup, fill = EthnicGroup), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar_parent + coord_polar()

```

```{r}
val_cnt <- as.data.frame(table(data_cleaned$EthnicGroup))
val_cnt_norm <- prop.table(table(data_cleaned$EthnicGroup)) * 100

val_cnt
val_cnt_norm <- as.data.frame(val_cnt_norm)
val_cnt_norm

mean_score_summary <- data_cleaned%>%
  group_by(EthnicGroup) %>%
  summarise_all(mean) %>% 
  select(EthnicGroup ,MathScore, ReadingScore, WritingScore)

mean_score_summary
```


Finding the correlation matrix between math, reading and writing scores. 

```{r}
scores_corr <- 
  data_cleaned %>%
  select(MathScore, ReadingScore, WritingScore) %>%
  data.frame() %>%
  janitor:: clean_names()
cor.matrix<-cor(cor(scores_corr, use="pairwise.complete.obs"))
corrplot(cor.matrix)
```


Lunch Type: 

- We notice that the amount of students on standard lunch is greater than the amount of student on free and reduced. 
- We can see that those on standard lunch have a higher correlation to having a higher mean score on all 3 sections of the test. 


```{r}
ggplot(data_cleaned, aes(x = LunchType)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()

```

```{r}
mean_scores_lunch <- data_cleaned %>%
  group_by(LunchType) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_lunch
```


Test Prep: 

- More people did not study for the test than those who did
- Those who did complete test prep have a higher mean in scores than those who didnt. 
- Associated with those who did not complete test prep could be bringing down the standard mean even further. 

```{r}
ggplot(data_cleaned, aes(x = TestPrep)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()
```



```{r}
mean_scores_prep <- data_cleaned %>%
  group_by(TestPrep) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_prep

```

Parent Marital Status: 

- Most parents seen are married within this school.
- Surprisingly those parents who are widowded and married had very similar test scores among the participants. 

```{r}
ggplot(data_cleaned, aes(x = ParentMaritalStatus)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()

bar_status <- ggplot(data = data_cleaned) + 
  geom_bar(
    mapping = aes(x = ParentMaritalStatus, fill = ParentMaritalStatus), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar_status + coord_polar()
```


```{r}
mean_scores_prep <- data_cleaned %>%
  group_by(ParentMaritalStatus) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_prep

```

Practice Sport: 

- More people play sports regularly or sometimes more than people who dont. 

- It appears to not have a massive affect on the outcome of the test scores

```{r}
ggplot(data_cleaned, aes(x = PracticeSport)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()

```

```{r}
mean_scores_prep <- data_cleaned %>%
  group_by(PracticeSport) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_prep

```

Is First Child:

- We can see that most students are first children in their family

- On average first children do better on the test then those who arent may be based on how much they are able to study. 


```{r}
ggplot(data_cleaned, aes(x = IsFirstChild)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()

```

```{r}
mean_scores_prep <- data_cleaned %>%
  group_by(IsFirstChild) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_prep

```

Number of siblings: 

-most people only have a few siblings

- since those with many siblings are a smaller proportion of the population, the data may be skewed based on that one significant persons intelligence/performance on test

```{r}
ggplot(data_cleaned, aes(x = NrSiblings)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()
bar_numb <- ggplot(data = data_cleaned) + 
  geom_bar(
    mapping = aes(x = NrSiblings, fill = NrSiblings), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar_numb + coord_polar()
```


```{r}
mean_scores_prep <- data_cleaned %>%
  group_by(NrSiblings) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_prep


```


Transport Means: 

- most people take the school bus
- no direct correlation on scores based on private versus public transit

```{r}
ggplot(data_cleaned, aes(x = TransportMeans)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()

```


```{r}
mean_scores_Trans <- data_cleaned %>%
  group_by(TransportMeans) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_Trans

```


Weekly Study Hours: 

- Most people wanted to study between 5- 10 hours

- No significant affect on if you study more than 10 hours compared to people who study between 5-10 hours

```{r}
ggplot(data_cleaned, aes(x = WklyStudyHours)) +
  geom_bar(fill = "aquamarine3", color = "black") +
  labs(title = "Gender") +
  theme_minimal()
bar_study <- ggplot(data = data_cleaned) + 
  geom_bar(
    mapping = aes(x = WklyStudyHours, fill = WklyStudyHours), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar_study + coord_polar()

```


```{r}
mean_scores_study <- data_cleaned %>%
  group_by(WklyStudyHours) %>%
  summarise(
    Mean_MathScore = mean(MathScore),
    Mean_ReadingScore = mean(ReadingScore),
    Mean_WritingScore = mean(WritingScore)
  )
mean_scores_study

```